CMPSubset

### 2021-08-20 CMP/CMPSubset.Rmd
Res1 dge analysis for biomark probes

@@

```{r setup}
library(dplyr)
library(Seurat)
library(patchwork)
library(ggplot2)
library(clustree)
```

# Set global variables
```{r}
projectName <- "CMP"
jackstraw.dim <- 40
```

store session info
```{r}
sessionInfo.filename <- paste0(projectName, "_sessionInfo.txt")
sink(sessionInfo.filename)
sessionInfo()
sink()
```

```{r}
source("../RFunctions/read_10XGenomics_data.R")
source("../RFunctions/PercentVariance.R")
source("../RFunctions/Mouse2Human_idconversion.R")
source ("../RFunctions/ColorPalette.R")
```

```{r warning=FALSE}
setwd("../../cellRanger/") # temporarily changing wd only works if you run the entire chunk at once
data_file.list <- read_10XGenomics_data(sample.list = "CMPm2")
object.data <-Read10X(data_file.list)
```

@@

@@

### Gene profiles of clusters
set ident at res = 1 and get markers
```{r}
Idents(cmp.object) <- "RNA_snn_res.1"
cmp.allmarkers.res1 <- FindAllMarkers(cmp.object)
cmp.top100markers.res1 <- cmp.allmarkers.res1 %>% group_by(cluster) %>% top_n(n = 100, wt = abs(avg_log2FC))
```

```{r}
cmp.top100markers.res1 <- cmp.allmarkers.res1 %>% group_by(cluster) %>% top_n(n = 100, wt = abs(avg_log2FC))
cmp.top100markers.res1 <- cmp.top100markers.res1[cmp.top100markers.res1$p_val_adj <= 0.05, ]
for(cluster in unique(cmp.top100markers.res1$cluster)){
	num.hits <- nrow(cmp.top100markers.res1[cmp.top100markers.res1$cluster == cluster, ])
	print(paste("cluster", cluster, "has", num.hits, "genes"))
}
```

@@

Here's the list of genes, plus some controls
```{r}
res1.biomark.genes <- c("Aif1",
"Aqp1",
"Birc5",
"Ccl3",
"Ccr2",
"Cd74",
"Cdc20",
"Cenpf",
"Cst3",
"Ctsh",
"Dntt",
"Elane",
"Ermap",
"F13a1",
"Fcer1g",
"Gm15915",
"Gm17590",
"H2-Aa",
"H2-Eb1",
"H2afx",
"H2afy",
"Hist1h2ac",
"Hmgb2",
"Hp",
"Ighm",
"Irf8",
"Lgals3",
"Lmo4",
"Ms4a2",
"Mt1",
"Plac8",
"Prtn3",
"Rap1b",
"Rgs1",
"Vwf",
"Wfdc17",
"Csrp3",
"Hist1h2ae",
"Ifitm1",
"Lgals1",
"Tmsb4x",
"Arl6ip1",
"Car2",
"Ccl9",
"Ccnb2",
"Cd9",
"Cenpa",
"Cpa3",
"Fos",
"Hist1h2ap",
"Ly6c2",
"Mpo",
"Pclaf",
"Slpi",
"Top2a",
"Ube2c",
"Ly86",
"Hist1h2bc",
"Pf4",
"Apoe",
"Ctsg",
"Car1",
"Hmmr", 
"Gapdh", 
"B2m", 
"Actb", 
"Hprt1",
"Gusb")
```

make vln plots per cluster
```{r}
for(gene in res1.biomark.genes){
	plot.title <- paste0(projectName, "dim", ndims, "res1_", gene)
	myplot<- VlnPlot(cmp.object, features = gene, pt.size = 0.01) + ggtitle(plot.title)
	png(filename = paste0(plot.title, ".png"), height = 800, width = 1600)
	plot(myplot)
	dev.off()

>>>>

## Update 2021.08.19
Pursuing Dim25 (85% of variance explained) res1 and res1.5 for biomark targets

@@

## Load libraries
```{r setup}
library(dplyr)
library(Seurat)
library(patchwork)
library(ggplot2)
library(clustree)
```

## Set global variables
```{r}
projectName <- "CMP"
jackstraw.dim <- 40
```

## Store session info
```{r}
sessionInfo.filename <- paste0(projectName, "_sessionInfo.txt")
sink(sessionInfo.filename)
sessionInfo()
sink()
```

## Load local scripts
```{r}
source("../RFunctions/read_10XGenomics_data.R")
source("../RFunctions/PercentVariance.R")
source("../RFunctions/Mouse2Human_idconversion.R")
source ("../RFunctions/ColorPalette.R")
```

# Read CMPm2
```{r warning=FALSE}
setwd("../../cellRanger/") # temporarily changing wd only works if you run the entire chunk at once
data_file.list <- read_10XGenomics_data(sample.list = "CMPm2")
object.data <-Read10X(data_file.list)
```

@@

### Clustree
what's the max resolution we can achieve while keepign clusters stable?
```{r fig.width = 15, fig.height = 5}
plot.title <- paste0(projectName, "_clustree_ndim", max(cmp.object@commands$RunUMAP.RNA.pca$dims))
my.clustree <- clustree(cmp.object, prefix = "RNA_snn_res.", node_colour = "sc3_stability", exprs = "scale.data") + 
	scale_color_continuous(low = 'red3', high = 'white') + 
	ggtitle(plot.title)
plot(my.clustree)
png(filename = paste0(plot.title, ".png"), height = 800, width = 1600)
plot(my.clustree)
dev.off()
```

@@

### Gene profiles of clusters
set ident at res = 1 and get markers
```{r}
Idents(cmp.object) <- "RNA_snn_res.1"
```

```{r}
cmp.allmarkers.res1 <- FindAllMarkers(cmp.object)
cmp.top100markers.res1 <- cmp.allmarkers.res1 %>% group_by(cluster) %>% top_n(n = 100, wt = abs(avg_log2FC))
```

```{r}
cmp.top100markers.res1 <- cmp.allmarkers.res1 %>% group_by(cluster) %>% top_n(n = 100, wt = abs(avg_log2FC))
cmp.top100markers.res1 <- cmp.top100markers.res1[cmp.top100markers.res1$p_val_adj <= 0.05, ]
for(cluster in sort(as.numeric(unique(cmp.top100markers.res1$cluster)))){
	num.hits <- nrow(cmp.top100markers.res1[cmp.top100markers.res1$cluster == cluster, ])
	print(paste("cluster", cluster, "has", num.hits, "genes"))
		cluster.markers <- FindMarkers(cmp.object, ident.1 = cluster)
		try(
			xlsx::write.xlsx(x = cluster.markers[,c("avg_log2FC", "p_val_adj")], 
											 file = paste0(projectName, "_dim", ndims, "_FindMarkersTop100.xlsx"), 
											 sheetName = paste0("clst", cluster), 
											 col.names = TRUE, 
											 row.names = TRUE, 
											 append = TRUE)
		)	
	}
```

@@

Here's the list of genes, plus some controls
```{r}
res1.biomark.genes <- c("Aif1",
												"Aqp1",
												"Birc5",
												"Ccl3",
												"Ccr2",
												"Cd74",
												"Cdc20",
												"Cenpf",
												"Cst3",
												"Ctsh",
												"Dntt",
												"Elane",
												"Ermap",
												"F13a1",
												"Fcer1g",
												"Gm15915",
												"Gm17590",
												"H2-Aa",
												"H2-Eb1",
												"H2afx",
												"H2afy",
												"Hist1h2ac",
												"Hmgb2",
												"Hp",
												"Ighm",
												"Irf8",
												"Lgals3",
												"Lmo4",
												"Ms4a2",
												"Mt1",
												"Plac8",
												"Prtn3",
												"Rap1b",
												"Rgs1",
												"Vwf",
												"Wfdc17",
												"Csrp3",
												"Hist1h2ae",
												"Ifitm1",
												"Lgals1",
												"Tmsb4x",
												"Arl6ip1",
												"Car2",
												"Ccl9",
												"Ccnb2",
												"Cd9",
												"Cenpa",
												"Cpa3",
												"Fos",
												"Hist1h2ap",
												"Ly6c2",
												"Mpo",
												"Pclaf",
												"Slpi",
												"Top2a",
												"Ube2c",
												"Ly86",
												"Hist1h2bc",
												"Pf4",
												"Apoe",
												"Ctsg",
												"Car1",
												"Hmmr")
hkgenes <- c("Gapdh", "B2m", "Hprt", "Pgk1", "Rplp2", "Pgk1", "Ubc", "Ywhaz", "Ppia", "Pum1", "Psmc4", "Elf1", "Mrpl19")
surface.markers<- c("Cd34", 
										"Kit",
										"Cd48", 
										"Ly6a",
										"Ly6e",
										"Cd9",
										"Itga2b",
										"Ly86",
										"Itga6",
										"Cd55",
										"Slamf1",
										"Flt3"
										)
```

make vln plots per cluster
```{r}
for(gene in res1.biomark.genes){
	plot.title <- paste0(projectName, "dim", ndims, "res1_", gene)
	myplot<- VlnPlot(cmp.object, features = gene, pt.size = 0.01) + ggtitle(plot.title)
	png(filename = paste0("VlnPlots/", plot.title, ".png"), height = 800, width = 1600)
	plot(myplot)
	dev.off()
}
for(gene in hkgenes){
	plot.title <- paste0(projectName, "dim", ndims, "res1_", gene)
	myplot<- VlnPlot(cmp.object, features = gene, pt.size = 0.01) + ggtitle(plot.title)
	png(filename = paste0("VlnPlots/", plot.title, "-hkgene.png"), height = 800, width = 1600)
	plot(myplot)
	dev.off()
}
for(gene in surface.markers){
	plot.title <- paste0(projectName, "dim", ndims, "res1_", gene)
	myplot<- VlnPlot(cmp.object, features = gene, pt.size = 0.01) + ggtitle(plot.title)
	png(filename = paste0("VlnPlots/", plot.title, "-surfMkr.png"), height = 800, width = 1600)
	plot(myplot)
	dev.off()
}
```

Plot candidate control genes
```{r fig.height=20, fig.width=10}
# VlnPlot(cmp.object, features = hkgenes, pt.size = 0.001, same.y.lims = FALSE, ncol = 2)
png(filename = paste0(projectName, "dim", ndims, "res1_TaqmanControls.png"), height = 1600, width = 2400)
VlnPlot(cmp.object, features = hkgenes, pt.size = 0.01, same.y.lims = FALSE, ncol = 5)
dev.off()
```
